Validity of a model for a cantilever two-dimensional beam

### 2024-04-07 files/Chapter5/PlateModel.tex
final small changes

\subsection{Galerkin approximation}
To discretise the body $\Omega$, the same shapes as in Section \ref{ssec:2D_Model:Galerkin} are used.

@@

Define the two spaces
\begin{eqnarray*}
	S^h_1 & = & \textrm{span}\left(\left\{\delta^1_i \ | \ i = 1,2,...,k_1 \right\} \right),\\
	S^h_2 & = & \textrm{span}\left(\left\{\delta^2_i \ | \ i = 1,2,...,k_2 \right\} \right).
\end{eqnarray*}
For each function $u^h \in S_1^h$ and each function $\psi^h \in S_2^h$, $u^h$ can be expressed as
\begin{eqnarray*}
	w^h = \sum_{i = 1}^{k} w_i(t) \delta^*_{i}(x)
\end{eqnarray*} and $\psi^h$ can be expressed as
\begin{eqnarray*}
	\psi^h = \sum_{i = 1}^{k} \psi_i(t) \delta_{i}(x)
\end{eqnarray*}

@@

The equation \eqref{Plate_FEM_M} is the same form as in Section \ref{2dFEM_EP} for the two-dimensional elastic body. Therefore the derivation of the eigenvalue problem is identical. Denote the eigenvalue problem for Problem P-1 by Problem P-1E.
\subsubsection{Problem P-1E}
Find a vector function $\bar{u}$ and a number $\lambda$ such that
\begin{eqnarray}
	M\lambda{\bar{u}} & = & K\bar{u}.
\end{eqnarray}

>>>>

\subsection{Galerkin approximation}
To discretise the body $\Omega$, the same shapes as in Section \ref{2d_FEM_G} are used.

@@

Define the two spaces
\begin{eqnarray*}
	S^h_1 & = & \textrm{span}\left(\left\{\delta^1_i \ | \ i = 1,2,...,k_1 \right\} \right),\\
	S^h_2 & = & \textrm{span}\left(\left\{\delta^2_i \ | \ i = 1,2,...,k_2 \right\} \right).
\end{eqnarray*}
For each function $u^h \in S_1^h$ and each function $\psi^h \in S_2^h$, $u^h$ can be expressed as
\begin{eqnarray*}
	w^h = \sum_{i = 1}^{k} w_i(t) \delta^1_{i}(x)
\end{eqnarray*} and $\psi^h$ can be expressed as
\begin{eqnarray*}
	\psi^h = \sum_{i = 1}^{k} \psi_i(t) \delta^2_{i}(x)
\end{eqnarray*}

@@

The equation \eqref{Plate_FEM_M} is in the same form as in Section \ref{2dFEM_EP} for the two-dimensional elastic body. Therefore the derivation of the eigenvalue problem is identical. Denote the eigenvalue problem for Problem P-1 by Problem P-1E.
\subsubsection{Problem P-1E}
Find a vector function $\bar{u}$ and a number $\lambda$ such that
\begin{eqnarray}
	M\lambda{\bar{u}} & = & K\bar{u}.
\end{eqnarray}

### 2024-04-07 files/Chapter6/Validity_2D.tex
final small changes

All of the results will include all the eigenvalues shared between the two-models, including not beam type eigenvalues. The non-beam type eigenvalues will be highlighted in grey. The non-beam type eigenvalues that are not shared between the two models will be excluded from the results, but the numbering of the eigenvalues will be kept as if they were included.

@@

Tables \ref{tab:2v3_1} and \ref{tab:2v3_2} show that the first 20 eigenvalues two-dimensional model compares very well to the matching eigenvalues three-dimensional beam for $b<=h$. As the width $b$ decreases, the two-dimensional model becomes a better approximation of the three-dimensional model. Also shown is that the slender beam with $h = 1/20$ compares better than the thick beam where $h = 1/5$, event though the case of $h=1/5$ is still a very good comparison.

@@

Tables \ref{tab:b>h} and \ref{tab:b>h_20} show that the two-dimensional model compares less well to the three-dimensional model when $b$ is too much larger than $h$. Tables \ref{tab:b>h-split} and \ref{tab:b>h-split_20} show that this is also true for the beam type eigenvalues.

@@

These results gives a detailed overview of the validity of the cantilever two-dimensional beam compared to the cantilever three-dimensional beam. The results show that the two-dimensional beam model compares very well to the three-dimensional beam model for a large range of beam shapes. The results shown were chosen to represent realistic cases, and the results are similar for other cases.

>>>>

All of the results will include all the eigenvalues shared between the two-models, including non-beam type eigenvalues. The non-beam type eigenvalues will be highlighted in grey. The non-beam type eigenvalues that are not shared between the two models will be excluded from the results, but the numbering of the eigenvalues will be kept as if they were included.

@@

Tables \ref{tab:2v3_1} and \ref{tab:2v3_2} show that the first 20 eigenvalues of the two-dimensional model compares very well to the matching eigenvalues of the three-dimensional beam for $b \leq h$. As the width $b$ decreases, the two-dimensional model becomes a better approximation of the three-dimensional model. Also shown is that the slender beam with $h = 1/20$ compares better than the thick beam where $h = 1/5$, event though the case of $h=1/5$ is still a very good comparison.

@@

Tables \ref{tab:b>h} and \ref{tab:b>h_20} show that the two-dimensional model compares not as well to the three-dimensional model when $b$ is larger than $h$. Tables \ref{tab:b>h-split} and \ref{tab:b>h-split_20} show that this is also true for the beam type eigenvalues.

@@

These results give a detailed overview of the validity of the cantilever two-dimensional beam compared to the cantilever three-dimensional beam. The results show that the two-dimensional beam model compares very well to the three-dimensional beam model for a large range of beam shapes. The results shown were chosen to represent realistic cases, and the results are similar for other cases.

### 2024-04-07 files/Chapter6/Validity_Plate.tex
final small changes

In the previous section, Section \ref{sec:validity-of-a-2d-beam}, it was shown that for certain applications beam models might not be the appropriate choice. The specific application is for models where the body has a larger width than height. A suggestion would be a plate model.

@@

Overall Tables \ref{tab:Table_plate_1}, \ref{tab:Table_plate_2} and \ref{tab:Table_plate_3} shows that the cantilever Reissner-Mindlin plate model compares very well to the three-dimensional plate. The maximum relative error is less than 3.65\% for all cases considered.

>>>>

In the previous section it was shown that for certain applications beam models might not be the appropriate choice. The specific application is for models where the body has a larger width than height. A suggestion would be a plate model.

@@

Overall Tables \ref{tab:Table_plate_1}, \ref{tab:Table_plate_2} and \ref{tab:Table_plate_3} show that the cantilever Reissner-Mindlin plate model compares very well to the three-dimensional plate. The maximum relative error is less than 3.65\% for all cases considered.

### 2024-04-07 files/Overview.tex
final small changes

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, three-dimensional elastic beam and plate models and a Reissner-Mindlin plate model. The aim in this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation are given (in dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. The variational forms are also derived which enables us to see the similarity between the models and the general theory.

@@

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. First idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed that follows the article \cite{CVV18}. Given a general vibration problem, it is split into two problems by introducing a trial solution. It can then be verified that these two problems are the eigenvalue problem and an ordinary differential equation. An additional assumption (additional to the assumptions of \cite{VV02}) is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has a complete orthonormal sequence of eigen-solutions. The solution of the general vibration problem is an infinite series of modal solutions. This formal series solution is then shown to also be valid for a initial value problem. This general case is also applicable to all the models in this dissertation.

@@

In real-world applications, a beam is a three-dimensional model. Therefore it should be more realistic to use a three-dimensional model to investigate the validity of the Timoshenko beam theory. This is mentioned by the authors of \cite{LVV09}. Their suggestion is to use the two-dimensional model as an intermediate step, to avoid complexities. Therefore the validity of the two-dimensional model is investigated, using a three-dimensional beam model as a reference. Again the results so that the comparison relies on the shape of the beams. The two-dimensional model compares well to the three-dimensional beam if the beam is not wide.

@@

The next result looks at modal analysis. Before the general case is discussed, again the cantilever Timoshenko beam is used to illustrated the concept of modal analysis.A trial solution to the boundary value problem is suggested. This trial solution is substituted into the partial differential equation and two problems are obtained. The first problem is the eigenvalue problem and the second is an ordinary differential equation. The eigenvalue problem can be solved with theory discussed later in the dissertation and the ordinary differential equation can then be solved. Substation of these two results into the boundary value problem confirms that the trial solution is correct. The same idea is then discussed for the general case.

>>>>

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, three-dimensional elastic beam and plate models and a Reissner-Mindlin plate model. The aim in this dissertation is to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exist, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation are given (in dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. The variational forms are also derived which enables us to see the similarity between the models and the general theory.

@@

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. First the idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed that follows the article \cite{CVV18}. Given a general vibration problem, it is split into two problems by introducing a trial solution. It can then be verified that these two problems are the eigenvalue problem and an ordinary differential equation. An additional assumption (additional to the assumptions of \cite{VV02}) is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has a complete orthonormal sequence of eigen-solutions. The solution of the general vibration problem is an infinite series of modal solutions. This formal series solution is then shown to also be valid for a initial value problem. This general case is also applicable to all the models in this dissertation.

@@

In real-world applications, a beam is a three-dimensional model. Therefore it should be more realistic to use a three-dimensional model to investigate the validity of the Timoshenko beam theory. This is mentioned by the authors of \cite{LVV09}. Their suggestion is to use the two-dimensional model as an intermediate step, to avoid complexities. Therefore the validity of the two-dimensional model is investigated, using a three-dimensional beam model as a reference. Again the results show that the comparison relies on the shape of the beams. The two-dimensional model compares well to the three-dimensional beam if the beam is not wide.

@@

The next result looks at modal analysis. Before the general case is discussed, again the cantilever Timoshenko beam is used to illustrated the concept of modal analysis.A trial solution to the boundary value problem is suggested. This trial solution is substituted into the partial differential equation and two problems are obtained. The first problem is the eigenvalue problem and the second is an ordinary differential equation. The eigenvalue problem can be solved with theory discussed later in the dissertation and the ordinary differential equation can then be solved. Substitution of these two results into the boundary value problem confirms that the trial solution is correct. The same idea is then discussed for the general case.
